Versioned data: why it’s needed and how it can be achieved (easily and cheaply)

### 2017-08-03 ms.tex
merging

Because \texttt{git} was built for tracking changes in code, there are aspects of how it behaves that are not natural to apply to  data \cite{Perkel-2016}.  For example, because it tracks changes line-by-line rather than cell-by-cell, there are inefficiencies in how the software stores the changes.  There is rapid current work on this problem and prototypes available, and as such, it is likely that \texttt{git} for data will replaced by a more natural system for data in the near future (REF).  This future system would be faster computationally, smaller in file size, and thus allow for larger datasets to be versioned efficiently.  That said, the workflow we describe here currently works well with \texttt{git} and should work more smoothly with its future replacement.

@@

TODO: turn this into a paragraph

@@

Tools for creating and archiving digital objects have seen rapid progress in recent years.  At present, each version of the dataset can be automatically archived and assigned a DOI, via several providers (Table \ref{tab:doi_minting}).  Here we describe the particular integration of github and zenodo.

@@

Using \texttt{datastorr}, researchers can set up their own versioned database (and R package) simply by providing: 1) a github repository name (e.g., ``traitecoevo/taxonlookup''); 2) a single file that will contain the versioned data (e.g., ``taxonlookup.csv''); 3) and the function used to load this into \texttt{R} (e.g., \texttt{read.csv()}). A tutorial explaining precisely how to set this up is available at \smurl{github.com/richfitz/datastorr}. % is this t

@@

The key features of a VDD system that are already possible are a versioning system to communicate small and large changes to users, both point-and-click and code-based access to all versions, reliable distribution to many users, and stable, organized DOIs.  Moreover, it is relatively easy to update and add new versions.  Technology is moving forward rapidly and so it is very likely that the particular tools we describe above will add features and be even easier to use.  But nonetheless, workable systems are already possible.

>>>>

Because \texttt{git} was built for tracking changes in code, there are aspects of how it behaves that are not natural to apply to  data \cite{Perkel-2016}.  For example, because it tracks changes line-by-line rather than cell-by-cell, there are inefficiencies in how the software stores the changes.  There is rapid current work on this problem and prototypes available, and as such, it is likely that \texttt{git} for data will replaced by a more natural system for data in the near future  \citep{Fli, Dat}.  This future system would be faster computationally, smaller in file size, flexible with respect to data structures, and thus allow for larger datasets to be versioned efficiently.  That said, the workflow we describe here currently works well with \texttt{git} and should work even more quickly with its future replacement.

@@

@@

Tools for creating and archiving digital objects have seen rapid progress in recent years.  At present, each version of the dataset can be automatically archived and assigned a DOI, via several providers (Table \ref{tab:doi_minting}).  Here we describe the particular integration of github and zenodo.  It is useful to have the ability to either be specific---reference a parti

@@

Using \texttt{datastorr}, researchers can set up their own versioned database (and R package) simply by providing: 1) a github repository name (e.g., ``traitecoevo/taxonlookup''); 2) a single file that will contain the versioned data (e.g., ``taxonlookup.csv''); 3) and the function used to load this into \texttt{R} (e.g., \texttt{read.csv()}). A tutorial explaining precisely how to set this up is available at \smurl{github.com/richfitz/datastorr}. % is this the right site?  ropensci?

@@

The key features of a VDD system that are already possible are a versioning system to communicate small and large changes to users, both point-and-click and code-based access to all versions, reliable distribution to many users, and stable, organized DOIs.  Moreover, it is relatively easy to update and add new versions.  Technology is moving forward rapidly and so it is very likely that the particular tools we describe above will add features and be even easier to use.  But, nonetheless, workable systems are already possible.
